---
title: "Project - Thera Bank"
author: "Juan Esteban Venegas"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

### 1. Project Objectives

The objective is to build a model for the marketing department of the Thera Bank to identify the potential customers who have a higher probability of purchasing a loan. This will increase the success ratio while at the same time reduce the cost of the campaign. 

The following steps will be performed throughout the project:

+ Perform exploratory data analysis on the dataset with the help of appropriate visualizations and identify observations/insights. 
+ Identify if there is evidence of multicollinearity in the dataset.
+ Perform simple linear regression for the dependent variable with every independent variable.
+ Perform PCA/Factor analysis by extracting 4 factors. 
+ Perform Multiple linear regression with customer satisfaction as dependent variables and the four factors as independent variables.

### 2. Exploratory Data Analysis – Step by step approach

#### 2.1 Environment Set up and Data Import

##### 2.1.1 Call necessary Packages and Invoke Libraries

```{r warning=FALSE, message=FALSE}
pacman::p_load(ggplot2, data.table, scales, corrplot, car, psych, broom, readxl, glmnet, nbClust, cluster, caret)
```

##### 2.1.2 Set up working Directory

```{r warning=FALSE, message=FALSE}
setwd('C:/Users/Juan Esteban Venegas/Desktop/Machine Learning Learning/Greatlearning/Project_4_Thera_Bank')
getwd()
```

##### 2.1.3 Import and Read the Dataset


```{r} Thera_Bank <- read_xlsx(paste0(getwd(),'/Thera Bank_Personal_Loan_Modelling-dataset-1.xlsx'), sheet = 'Bank_Personal_Loan_Modelling')```
```{r}
ReadMe <- as.data.table(read_xlsx(paste0(getwd(),'/Thera Bank_Personal_Loan_Modelling-dataset-1.xlsx'), sheet = 'ReadMe'))
names(ReadMe) <- c('Variable', 'Description')
ReadMe <- ReadMe[!is.na(Variable)]
```

```{r}
Thera_Bank <- as.data.table(read_xlsx(paste0(getwd(),'/Thera Bank_Personal_Loan_Modelling-dataset-1.xlsx'), sheet = 'Bank_Personal_Loan_Modelling'))
names(Thera_Bank) <- gsub("\\.","",make.names(gsub("(?<=[\\s])\\s*|^\\s+|\\s+|\\.$", "_", names(Thera_Bank), perl=TRUE)))
```

#### 2.2 Exploratory data analysis

##### 2.2.1 Variable Identification – Inferences

```{r}
names(Thera_Bank)
```

There are 14 columns in the data set with `r nrow(Thera_Bank)` observations.

```{r}
str(Thera_Bank)
```

All variables are numeric but some need to be transformed to factor.
```{r}
Thera_Bank[,c('Education', 'Personal_Loan', 'Securities_Account', 'CD_Account', 'Online', 'CreditCard', 'ZIP_Code')] <- lapply(Thera_Bank[,c('Education', 'Personal_Loan', 'Securities_Account', 'CD_Account', 'Online', 'CreditCard', 'ZIP_Code')] , factor)
```

```{r}
str(Thera_Bank)
```

```{r}
dmy <- dummyVars(" ~ .", data = Thera_Bank)
Thera_Bank <- data.frame(predict(dmy, newdata = Thera_Bank))
head(Thera_Bank,4)
```
There are `r nrow(Thera_Bank[!complete.cases(Thera_Bank)])`) missing values. All of these correspond to Family_member information. Because the number of customers without this information is so low and the value is hard to properly inpute, these values will be excluded from the model for the time being.

```{r}
summary(Thera_Bank)
Thera_Bank_nonull <- Thera_Bank[!is.na(Family_members)]
```


```{r}
par(mfrow = c(1,6))
boxplot(Thera_Bank_nonull$Age_in_years, data = Thera_Bank_nonull, ylab = 'Age_in_years')
boxplot(Thera_Bank_nonull$Experience_in_years, data = Thera_Bank_nonull, ylab = 'Experience_in_years')
boxplot(Thera_Bank_nonull$Income_in_Kmonth, data = Thera_Bank_nonull, ylab = 'Income_in_Kmonth')
boxplot(Thera_Bank_nonull$Family_members, data = Thera_Bank_nonull, ylab = 'Family_members')
boxplot(Thera_Bank_nonull$CCAvg, data = Thera_Bank_nonull, ylab = 'CCAvg')
boxplot(Thera_Bank_nonull$Mortgage, data = Thera_Bank_nonull, ylab = 'Mortgage')

```

#### 2.3 Univariate Analysis

```{r warning=FALSE, message=FALSE , fig.width=14, fig.height=12}
ggplot(melt(Thera_Bank_nonull[,c(1:4,6,7,9)], id.vars = 'ID', variable.name = 'Variable', value.name = 'Value'), aes(x = Value)) + geom_density() + facet_grid(.~Variable, scales = 'free') + labs(title = 'Density Plot for continuous variables', x = 'Value', y = 'Frequency') + theme(axis.text.x = element_text(angle = 90))
```

Age and experience seem to have similar distributions with the sample spread evenly. Income, CCAvg and Mortgage seem to be right skewed.

#### 2.4 Bi-Variate Analysis

```{r warning=FALSE, message=FALSE , fig.width=12, fig.height=12}
cormatrix <- cor(Thera_Bank_nonull[,c(2:4,6,7,9)])
corrplot.mixed(cormatrix, upper ='square', lower.col = 'black')
```
When looking at correlations between numeric variables, we can see that the strongest correlation between independent variables is Age and Experience. There is also a positive correlation between income and average spending on credit cards per month. 


### 3. Clustering

#### 3.1 Kmeans Clustering

```{r}
set.seed(1000) 
Thera_Bank_nonull_scaled <- scale(Thera_Bank_nonull[,c(2:4,6,7,9)])
nc <- NbClust(Thera_Bank_nonull_scaled, min.nc=2, max.nc=5, method="kmeans")
```

Kmeans Clustering works with numeric data because it uses distance between observations to determine if they are similar or not. Because of this, the first step is to select numeric data from the table and scale it. We then run NbClust to determine what the best number of clusters is for this data set.


```{r warning=FALSE, message=FALSE}
Kmeans_clusters <- kmeans(Thera_Bank_nonull_scaled, centers = 3, nstart = 5)
clusplot(Thera_Bank_nonull_scaled, Kmeans_clusters$cluster, color=TRUE, shade=FALSE, labels = 0, col.p = Kmeans_clusters$cluster, stand = TRUE)

```


```{r warning=FALSE, message=FALSE}
```


```{r warning=FALSE, message=FALSE}
```

#### 3.1 EFA


```{r warning=FALSE, message=FALSE}
```

```{r warning=FALSE, message=FALSE}
```

```{r warning=FALSE, message=FALSE}
```


```{r warning=FALSE, message=FALSE}
```


```{r warning=FALSE, message=FALSE}
```


```{r warning=FALSE, message=FALSE}
```


```{r warning=FALSE, message=FALSE}
```